%-----------------------------------------------------------------
\begin{plSection}{Foundations of mathematics}
\label{sec:Foundations_of_mathematics}

\begin{plQuote}
{\citeAuthorYearTitle{Thurston:1994:Proof}}
{}
On the most fundamental level, the foundations of mathematics are much shakier
than the mathematics that we do. Most mathematicians adhere to foundational
principles that are known to be polite fictions. For example, it is a theorem that
there does not exist any way to ever actually construct or even define a well-ordering
of the real numbers. There is considerable evidence (but no proof) that we can get
away with these polite fictions without being caught out, but that doesn’t make
them right. Set theorists construct many alternate and mutually contradictory
``mathematical universes'' such that if one is consistent, the others are too. This
leaves very little confidence that one or the other is the right choice or the natural
choice. {\Godel}’s incompleteness theorem implies that there can be no formal system
that is consistent, yet powerful enough to serve as a basis for all of the mathematics
that we do.
\end{plQuote}

\begin{plQuote}
{Dedekind to Klein, April 1888, 
from \\
\citeAuthorYearTitle{Dugac:1976:DedekindFondements},\\
as quoted in \\
\citeAuthorYearTitle{Ferreiros:2007:Labyrinth}.}
{}
{Und was wird der geduldige Leser am
Schlusse sagen? Dass der Verfasser mit einem Aufwande von uns/iglicher Arbeit es gliicklich
erreicht hat, die klarsten Vorstellungen in ein unheimliches Dunkel zu hiillen!}
\par
What will the forbearing reader say at the end? That the author,
 in a squandering of indescribable work, 
 has happily managed to surround 
the clearest ideas in a disturbing obscurity!
\end{plQuote}

%-----------------------------------------------------------------
\begin{plSection}{Crisis 1870--1930}
\label{sec:Crisis-1870--1930}
%-----------------------------------------------------------------
\begin{plSection}{Ferreir\'{o}s Crisis}
\label{sec:Ferreiros_Crisis}

Quote and paraphrase from 
\citeAuthorYearTitle{Ferreiros:2008:Crisis}.

G\"{o}ttingen/modern 
(Gauss, Dirichlet, Riemann, Dedekind, Cantor, Hilbert)
vs
Berlin, Paris (Kronecker, Weierstrauss)

Dissents from modern approach by Cantor and Riemann.

'Modern' approach:
\begin{itemize}
\item Dirichelet's 'arbitrary' functions
\item infinite sets and higher infinities
\item ``thoughts in place of calculations'' (Dirichelet)
\item structures characterized axiomatically.
\item purely existential methods of proof.
\end{itemize}

Early example: Dedekind (1871) Algebraic Number Theory,
Number Fields. Ideals.
 (``An ideal I in a ring R is a subset 
 such that the sum and difference of any two elements of I 
 and the product of any element of I with any element of R 
 are also in I.~\cite{sep:DedekindFoundations})
Existential proof of unique prime factorization for
ideals in any ring jof algebraic integers.
Kronecker complains about ability to \textsl{calculate}
(or \textsl{construct})
either ideals of their factors.
Dedekind: concrete problems require 
'more delicate computational techniques'~\cite{Ferreiros:2008:Crisis},
but conceptual theory is important.
Note similarity to Riemann's treatment of $\Space{C}$.
1867--1872

Wierestrauss: analytic functions via explicit power series;
Riemann: analytic functions satisfy differentiability condition.
$\Rightarrow$ Weierstrauss examples of nowhere differentiable
continuous functions.

Weierstrauss: function as analytical expression 
(how close is this to function as procedure?);
Riemann, Dedekind, Direchelet: function as single-valued relatio.
'Right framework'' for continuity and integration.
Called the \textit{conceptual approach} in 1800s math.

Kronecker disbelief in Bolzano-Weierstrauss thm.
(every infinite bounded set in $\Space{R}$ 
has an accumulation point);
accumulation pt can't be constructed from $\Space{Q}$.

Cantor's proofs in set theory ``quintessential examples
of modern methodology of existential proof''; 
attacks on Kronecker 1883.
Kronecker 1887 paper on foundational views.

Hilbert: axiomatic method in geometry 1899.

Cantor, Russell, Zermelo: circa 1900, paradoxes:
\begin{itemize}
  \item set theoretic: assumption of existence of certain sets
  leads to contradiction.
  \item semantic: truth and definability
\end{itemize}

\end{plSection}%{Ferreir\'{o}s Crisis}
%-----------------------------------------------------------------
\end{plSection}%{Crisis 1870--1930}
%-----------------------------------------------------------------
%-----------------------------------------------------------------
\input{feferman}
%-----------------------------------------------------------------
\begin{plSection}{Constructivism}
\label{sec:Constructivism}
%-----------------------------------------------------------------
\begin{plSection}{SEP}
\label{sec:Constructivism_SEP}
\cite{sep:ConstructiveMathematics}

Multiple references to ``the informal, though rigorous, 
style of the practicing analyst, algebraist, topologist, \ldots''
???

Section 1 has problems:

Discussion of
$ \forall x \in \Space{R} (x = 0 \vee x \ne 0)$.

``However, because the computer can handle real numbers 
only by means of finite rational approximations, 
we have the problem of underflow, 
in which a sufficiently small positive number can be misread as 0 
by the computer;\ldots''

\ldots which is nonsense;
real issue is non-termination of test for $x=0$.

Section 2:

BHK interpretation~\cite{wiki:BrouwerHeytingKolmogorovInterpretation}:
\begin{itemize}
  \item A proof of $P\wedge Q$ 
  is a pair $(a,b)$ where $a$ is a proof of $P$ 
  and $b$ is a proof of $Q$.
\item A proof of $P\vee Q$ is a pair $(a,b)$ 
where $a$ is $0$ and 
$b$ is a proof of $P$, 
or $a$ is $1$ 
and $b$ is a proof of $Q$.
\item A proof of $P\Rightarrow Q$ is 
a function $f$ that converts 
a proof of $P$ into a proof of $Q$.
\item A proof of $\exists x\in S:\varphi (x)$ is 
a pair $(a,b)$ where 
$a$ is an element of $S$, 
and $b$ is a proof of $\varphi (a)$.
\item A proof of $\forall x\in S:\varphi (x)$ is 
a function $f$ that converts 
an element (any element?) $a$ of $S$ into a proof of $\varphi (a)$.
\item The formula $\neg P$ is defined as $P\Rightarrow \bot$, 
so a proof of it is a function $f$ that converts 
a proof of $P$ into a proof of $\bot$.
\item There is no proof of $\bot$
(the absurdity, or bottom type 
(non-termination in some programming languages)).
\end{itemize}
What is a 'function' here? 
Other sources use 'algorithm', also without definition.

Analysis examples too detailed and lacking motivation 
to make point clear, feel cut-and-pasted from the middle
of a more coherent presentation:

Bishop: constructive $\Space{R}$ 
from Cauchy sequences 
in $\Space{Q}$~\cite{BishopBridges:1985:ConstructiveAnalysis}.

Bridges and Vita: constructive $\Space{R}$ from sequences
of intervals in $\Space{Q}$~\cite{Bridger:2019}.

Intermediate value theorem C: 
``If $f$ is a continuous real-valued mapping 
on the closed interval $[0,1]$ such that $f(0)⋅f(1)<0$, 
then there exists $x$ such that $0<x<1$ and $f(x)=0$.

Proof requirement: 
An algorithm which, applied to the function $f$, 
a modulus of continuity for $f$, and the values $f(0)$ and $f(1)$,
computes an object $x$ and shows that $x$ is a real number 
between $0$ and $1$; and
shows that $f(x)=0$.''~\cite{sep:ConstructiveMathematics}

Not constructive because it requires
``lesser limited principle of omniscience (LLPO)''.

Intermediate value theorem D:
``If $f$ is a continuous real-valued mapping 
on the closed interval $[0,1]$ 
such that $f(0)⋅f(1)<0$, 
then for each $\epsilon>0$ 
there exists $x$ such that $0<x<1$ and $|f(x)|<\epsilon$.
Proof requirement: 
An algorithm which, applied to the function $f$, 
a modulus of continuity for $f$, the values $f(0)$ and $f(1)$, 
and a positive number $\epsilon$,
computes an object $x$ 
and shows that $x$ is a real number between $0$ and $1$; and
shows that $|f(x)|<\epsilon$.''~\cite{sep:ConstructiveMathematics}

Can be proved constructively.

Stronger intermediate value theorem:
``Let $f$ be a continuous real-valued mapping 
on the closed interval $[0,1]$ 
such that $f(0)⋅f(1)<0$. 
Suppose also that $f$ is locally nonzero, 
in the sense that for each $x \in [0,1]$
and each $r>0$, 
there exists $y$ such that $|x−y|<r$ and $f(y) \neq 0$. 
Then there exists $x$ such that $0<x<1$ and $f(x)=0$.''

Section 3: Varieties

Section 3.1: Intuitionist (Brouwer)

``Brouwer was not the clearest expositor of his ideas, 
as is shown by the following quotation:

Mathematics arises when the subject of two-ness, 
which results from the passage of time, 
is abstracted from all special occurrences. 
The remaining empty form [the relation of n to n+1] 
of the common content of all these two-nesses 
becomes the original intuition of mathematics 
and repeated unlimitedly creates new mathematical subjects. 
(quoted in Kline [1972], 1199–2000)''

(Nederlands to Deutsch to English?)

``According to [Brouwer’s] view and reading of history, 
classical logic was abstracted 
from the mathematics of finite sets and their subsets. \ldots 
Forgetful of this limited origin, 
one afterwards mistook that logic 
for something above and prior to all mathematics, 
and finally applied it, without justification, 
to the mathematics of infinite sets. 
This is the Fall and original sin of set theory, 
for which it is justly punished by the antimonies. 
It is not that such contradictions showed up that is surprising,
 but that they showed up at such a late stage of the game. 
 (Weyl [1946])''
 
Weyl may have worked out a description of
$\Space{R}$ without (uncountable) infinity or
axiom of choice?

\end{plSection}%{SEP}
%-----------------------------------------------------------------
\begin{plSection}{IEP}
\label{sec:Constructivism_IEP}
\cite{iep:ConstructiveMathematics}

``An implication ($A \Rightarrow B$) is not equivalent 
to a disjunction ($\lnot A \vee B$), 
and neither are equivalent to a negated conjunction 
($\not (A \wedge \lnot B$)). 
In practice, constructive mathematics may be viewed 
as mathematics done using intuitionistic 
logic.''\cite{iep:ConstructiveMathematics}

Hilbert vs Brouwer circa 1900: 
non-constructive existence theorems via contradiction.

Usual example: Prove $\exists a,b \notin \Space{Q}$
such that $a^b \in \Space{Q}$. \hfill\break
Non-constructive: Use law of excluded middle.
If $\sqrt{2}^{\sqrt{2}} \in \Space{Q}$, done.
If not, consider ${\sqrt{2}^{\sqrt{2}}}^{\sqrt{2}}$.
\hfill\break
Constructive: consider $a=\sqrt{2}$ and $b=\log_2(9)$.

BHK interpretation of intuitionistic 
logic\cite{wiki:BrouwerHeytingKolmogorovInterpretation}.

\end{plSection}%{IEP}
%-----------------------------------------------------------------
\begin{plSection}{Martin-L\"{o}f}
\label{sec:Martin_Lof_IEP}

``However he soon revisited foundations in a different way, 
which harks back to Russell's type theory, 
albeit using a more constructive and less logicist approach. 
The basic idea of Martin-Löf's theory of types (Martin-Löf, 1975) 
is
that mathematical objects all come as types, 
and are always given in terms of a
type (for example, one such type is that of functions 
from natural numbers to
natural numbers), due to an intuitive understanding 
that we have of the notion
of the given type.

The central distinction Martin-Löf makes is that 
between proof and derivation.
Derivations convince us of the truth of a statement, 
whereas a proof contains the data necessary for computational
(that is, mechanical) verification of a proposition. 
Thus what one finds in standard mathematical textbooks are
derivations; 
a proof is a kind of realizability (c.f. Kleene, 1945) and links
mathematics to implementation (at least implicitly).

The theory is very reminiscent of a kind of cumulative hierarchy. 
Propositions can be represented as types 
(a proposition's type is the type of its proofs),
and to each type one may associate a proposition 
(that the associated type is not empty). 
One then builds further types by construction on already existing
types.''

\end{plSection}%{Martin-L\"{o}f}
%-----------------------------------------------------------------
\begin{plSection}{Constructivism (Wikipedia)}
\label{sec:Constructivism_Wikipedia}

\cite{wiki:ConstructivismPhilosophyOfMathematics}

Usually drop law of excluded middle 
($P \vee \lnot P$).\cite{wiki:LawOfExcludedMiddle}
It is possible that both $P$ and $\lnot P$ are not provable,
so $P \vee \lnot P$ is not provable.

But keep law of non-contradiction
($\lnot ( P \wedge (\lnot P))$).~\cite{wiki:LawOfNoncontradiction}
In other words, it is impossible to prove both.

Witnesses: Proving $\exists _{x\in \Set{X}}P(x)$ requires
constructing a specific $a \in \Set{X}$ such that $P(a)$ is true; 
$a$ is a \textit{witness} for $\exists _{x\in \Set{X}}P(x)$.

People\cite{wiki:ConstructivismPhilosophyOfMathematics}:
\begin{description}
\item[Leopold Kronecker] old constructivism, semi-intuitionism.
\item[Luitzen Egbertus Jan (Bertus) Brouwer] forefather of intuitionism.
\item[Andrey Andreyevich Markov, Jr. ] forefather of Russian school of constructivism.
\item[Arend Heyting] formalized intuitionistic logic and theories.
\item[Per Martin-Löf] founder of constructive type theories.
\item[Errett Bishop] promoted a version of constructivism 
claimed to be consistent with classical mathematics.
\item[Paul Lorenzen] developed constructive analysis.
\end{description}

\end{plSection}%{Constructivism (Wikipedia)}
%-----------------------------------------------------------------
\begin{plSection}{Constructive analysis}
\label{Constructive analysis}

In constructive 
analysis~\cite{wiki:ConstructiveAnalysis,
Bridger:2019,Henle:2012:RealNumbers},
varieties of constructionism are divided by
what sort of countable $\rightarrow$ countable functions can be 
constructed?~\cite{wiki:ConstructivismPhilosophyOfMathematics}
\begin{description}
\item[Free choice sequences] 
\cref{sec:Choice_sequences}\cite{wiki:ChoiceSequence}.
\item[Algorithms/Computable functions] 
\cref{sec:Computable_function}~\cite{wiki:ComputableFunction}
 results in computable numbers (reals?).
\end{description}

\end{plSection}%{Constructive analysis}
%-----------------------------------------------------------------
\begin{plSection}{Choice sequences}
\label{sec:Choice_sequences}

\begin{description}
\item[Law-like] determined completely (somehow?),
eg, $\Space{N}$ or functions $\Space{N} \mapsto \Space{N}$,
functions from the first $k$ elements of the sequence to 
$\Space{N}$.
(How is this different from a computable function?)
\item[Lawless] only a finite prefix determined at any point.
Example is rolls of a die. (Very dubious\ldots.)
\end{description}

Axioms 
(Are these only for lawless sequences?
They don't seem 'constructive' to me\ldots):
\begin{description}
\item[open data] Let $\alpha \in n$ mean that the first items
in the choice sequence $\alpha$ match the finite sequence $n$.
Then for any predicate $A$,
$A(\alpha )$ implies  $\exists n$ such that
$\alpha \in n$ and $\forall \beta \in n$ we have $A(\beta )$.
In other words, the truth of any predicate on a sequence can be
determined from a finite prefix.
\item[density] $\forall n\,\exists \alpha [\alpha \in n]$.
There exists some sequence that has any given finite prefix.
\end{description}

\end{plSection}%{Choice sequences}
%-----------------------------------------------------------------
\begin{plSection}{Computable function}
\label{sec:Computable_function}
\cite{wiki:ComputableFunction}

\end{plSection}%{Computable function}
%-----------------------------------------------------------------
\begin{plSection}{Cardinality of \texorpdfstring{$\Space{R}$}{R}}
\label{sec:Cardinality_of_R}

Issues with Cantor's diagonal 
argument~\cite{wiki:CantorsDiagonalArgument}.

(Starts with a given enumeration of (say) binary expansions of 
elements of $\Space{R}$, then \textit{constructs} a missing
element. But where does the enumeration come from?
Seems like a proof by contradiction\ldots?)

\end{plSection}%{Cardinality of $\Space{R}$}
%-----------------------------------------------------------------
\begin{plSection}{Axiom of choice}
\label{sec:AxiomOfchoice}

One interpretation of 'constructive' is
'provable in $\textsf{ZF}$ set theory without the axiom of 
choice.'~\cite{iep:SetTheory,wiki:AxiomOfChoice}
(See \cref{sec:Zermelo-Fraenkel-set-theory} for \textsf{ZF}.)
But \textsf{ZF} may be considered 'not constructive'
itself.

In some set theories, the axiom of choice implies
the law of excluded middle; in others it doesn't.
When it does, constructionists use weaker versions
that don't imply $P \, \wedge \, \lnot P$.
So some constructionists restrict themselves

\end{plSection}%{Axiom of choice}
%-----------------------------------------------------------------
\begin{plSection}{Measure theory}
\label{sec:Measure_theory}

\citeAuthorYearTitle{Bishop:1967:Foundations,BishopCheng:1972:ConstructiveMeasureTheory,BishopBridges:1985:ConstructiveAnalysis}

\end{plSection}%{Measure theory}
%-----------------------------------------------------------------
\end{plSection}%{Constructivism}
%-----------------------------------------------------------------
\end{plSection}%{Foundations of mathematics}
 